Article Title

### 2018-03-09 doc/turtles_plosone.tex
combined cc7g and tracg so it behaves

\afterpage{\clearpage}
\begin{figure}[ht]
  \centering
  \includegraphics[height = 0.8\textheight, width = \textwidth, keepaspectratio = true]{figure/Ken_Ang_SpLoc_final_comp}
  \caption{Geographic distribution of specimens sampled for comparing the hypothesized subdivisions of \textit{Emys marmorata}. Each hypothesized scheme has two or more possible classes. Sample size differs between schemes because of our ability to confidently assign museum specimens to the various schemes. The number of localities shown on each map is less than the number of specimens sampled because some localities produced multiple specimens. The different classification abbreviations are as follows: \textit{E. marmorata} = ``marm'', \textit{E. pallida} = ``pall'', Central Coast Ranges = ``CCR'', San Joaquin Valley = ``SJ,'' Baya Peninsula = ``Baja,'', and Sierra Foothills = ``Foothill.''}
  \label{fig:map}
\end{figure}

\afterpage{\clearpage}
\begin{figure}[ht]
  \centering
  \includegraphics[height = 0.5\textheight, width = \textwidth, keepaspectratio = true]{figure/plastra}
  \caption{Depiction of general plastral shape of \textit{E. marmorata} and position of the 19 landmarks used in this study. Anterior is towards the top of the figure.}
  \label{fig:plastra}
\end{figure}

Three different geometric morphometric datasets describing turtle plastron variation were assembled for this analysis: 1) specimens from seven distinct emydine species and a deriochelyine outgroup (\textit{C. picta}); 2) \textit{T. scripta} specimens from the two main subspecies (\textit{T. scripta elegans} and \textit{T. scripta scripta}); and 3) \textit{E. marmorata} specimens from across the species' geographic range. The first two datasets are intended to serve as a test of whether machine learning techniques can differentiate species-level groupings of emydid turtles using plastron shape. We expect that the first case represents a low complexity dataset because of the high level of plastron shape disparity that exists among these species \cite{Claude2003a,Claude2006,Angielczyk2011}, whereas the second dataset should be relatively higher in complexity and more analogous to the \textit{E. marmorata} example. We predict that the \textit{E. marmorata} dataset should be of the highest complexity and our greatest challenge given the previous finding that only very subtle differences existed between geographically-distinct populations \cite{Holland1992}.
%We expect the first dataset represents a low-complexity dataset given the high level of plastron shape disparty that exists among these species \cite{Angielczyk2011}, whereas the second dataset should be relatively higher in complexity than the first. Ultimately, we expect that the third dataset to be of the highest complexity and our biggest analytical challenge. For all datasets we chose to focus on adults because significant changes in plastron shape occur over the course of ontogeny in \textit{E. marmorata} and other emydines \cite{Angielczyk2013a}.
% three turtle datasets
%   eight species
%   Trachemys species
%   Emys complex
% 26 landmarks

@@

@@

Sex was known only for a subset of the total dataset and was not included as a predictor of classification. Instead, we estimated the degree to which specimens cluster morphologically by sex in order to determine how much of a potential biasing factor sexual dimorphism could be for our analysis of the \textit{E. marmorata} species complex (see below).
%The scheme names are as follows: Mito 1 and 2 correspond to \cite{Spinks2005}, Mito 3 corresponds to \cite{Spinks2010}, Morph 1 and Morph 2 correspond to \cite{Holland1992}, and Nuclear corresponds to \cite{Spinks2014}. % CHANGE

@@

@@

@@

Instead of relying on a single supervised learning method, we chose to use an ensemble approach where multiple model types are used in concert so that any congruence between them increases our support for that conclusion over another \cite{Hastie2009}. The supervised learning methods used here are named in Table \ref{tab:methods}. Each of these methods makes different assumptions, treats data differently, and can produce different classification results depending on the nature of the data \cite{Hastie2009}. For example, multinomial logistic regression is a type of generalized linear model, whereas random forest is itself an ensemble approach where multiple decision trees are fit to subsets of the full dataset and then averaged.
% use multiple methods
%   multinomial logistic regression (nnet)
%   linear discriminate analysis (MASS)
%   penalized discriminate analysis (mda)
%   neural network (nnet)
%   random forest (randomForest)
% each method has different assumptions and treat data differently
%   all assume that predictors have additive effect (e.g. independent)

% predictors/features/covariates
%   1:25 PCs
%   size
%   size X PC1
% analogy to PCA regression
The maximum set of possible predictors or features used for any model of our dataset is comprised of the first 25 principal components (PCs), scaled centroid size, the interaction between scaled centroid size and PC 1, and the interaction between scaled centroid size and PC 2. Additional interaction terms were not considered because of model complexity/sample size concerns. Size and the interaction between size and PCs 1 and 2 were included as predictors to account for known ontogenetic variation in plastron shape \cite{Angielczyk2013a} as well as potential size differences between classes, even if this is unlikely \cite{Seeliger1945,Holland1992}. These data constitute a ``maximum set'' because the best or selected models based on five-fold cross-validation need not, and likely will not, include all predictors possible (see below). Because our supervised learning models use PCs as predictors, this approach is in many ways analogous to PCA regression. PCA regression takes advantage of the reduction and orthogonality PCs to improve regression fit \cite{Hastie2009}. Because the PCs of shape are by definition orthogonal, they can easily serve as independent predictors or features of class membership without fear of collinearity.

% training and testing paradigm
%   training dataset
%     for multiple models with between 3 and 28 predictors
%     5-fold cross validation of each model
%     best model had greatest mean AUC value
%     selected model is most parsimonious model within 1 SD of ROC of best model
%   testing dataset
%     estimate class of out-of-sample testing dataset
%     compare average AUC across models for each scheme
% caret package for R
We adopted a training and testing paradigm for selecting parsimonious models and estimating their overall error rates \cite{Hastie2009,Kuhn2013}. Within-sample model performance is inherently biased upwards, so model evaluation requires overcoming this bias. With very large sample sizes, as in this study, part of the sample can be used as the ``training set'' and the remainder acts as the ``testing set.'' In this approach, following all cleaning and vetting, the data are split into a training dataset and a testing dataset. The former is used for fitting the model whereas the later is used for measuring model performance, a process called model generalization. For each scheme, we limited the model training and testing to only those individuals with class labels for that scheme. In this analysis, we randomly divided 80\% of samples into the training set and the remaining 20\% into the testing set.

% auc of roc as performance measure
%   relationship between false positive rate and true positive rate
%   varies between 0.5 and 1 (random -- perfect)
%   better performance than accuracy when groups unbalanced
In classification studies, such as this one, a common metric of performance is the receiver operating characteristic (ROC) which is the relationship between the false and true positive rates \cite{Hastie2009}. The area under the ROC curve (AUC) is the derived estimate of model performance; AUC ranges from 0.5 to 1 which correspond to performance similar to random guesses and perfect classification rates, respectively \cite{Hastie2009}. Both ROC and AUC are preferable to simple classification accuracy when class membership is unbalanced, as it is in these analyses \cite{Hastie2009}. The standard ROC and AUC calculations are defined only for binary classifications, which is not the case for our eight species and \textit{Emys} complex datasets. To generalize this approach for situations with multiple response classes, we used an all-against-one strategy where the model AUC is the average of the AUC values from the multiple binary comparisons of one class compared to all others \cite{Hand2001}.

@@

\begin{figure}[h]
  \centering
  \begin{subfigure}[b]{0.7\textwidth}
    \caption{}
    \includegraphics[width = \textwidth]{figure/cc7_pc_graph}
  \end{subfigure}

  \begin{subfigure}[b]{0.7\textwidth}
    \caption{}
    \includegraphics[width = \textwidth]{figure/tra_pc_graph}
  \end{subfigure}
  \caption{Two scatterplots of morphological differences from two of the three datasets analyzed in this study. (a) Scatterplot of the first two PCA axies from the landmarks from the eight different species dataset, and (b) the first two axes of variation from two subspecies of \textit{Trachemys} dataset. Point colors correspond to the categories within each dataset while point size is proportional to individual centroid size. In parentheses next to the axis labels are the percent of total variation accounted for by that axis. For both datasets there are clear distinctions between the different categories.}
  \label{fig:other_pca}
\end{figure}

\afterpage{\clearpage}
\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/emys_pc_graph}
  \caption{Scatterplot of the first two axes of morphological variation in the \textit{Emys marmorata} dataset. Each panel corresponds to one of the six different classification schemes analyzed as part of this study (Tab. \ref{tab:hypotheses}). Point color corresponds to the categories within each scheme, and the class names correspond to geographic regions. Point size is proportional to centroid size of that specimen and the numbers in parentheses next to the axis labels are the percent of total variation accounted for along that axis.}
  \label{fig:emys_pca}
\end{figure}

@@

@@

\begin{figure}[h]
  \centering
  \includegraphics[height = 0.3\textheight, width = \textwidth, keepaspectratio = true]{figure/sex_test_hist}
  \caption{Comparison of observed Procrustes distance between the centroids of each sex (vertical line) to a null distribution generated from 1000 permutations of the sex-labels. This result indicates that the difference between the centroids is as small/smaller than expected by random.}
  \label{fig:sex_test}
\end{figure}

\afterpage{\clearpage}
\begin{figure}[h]
  \centering
  \includegraphics[height = 0.8\textheight, width = \textwidth, keepaspectratio = true]{figure/sex_test_hist_grouped}
  \caption{Comparison of observed Procrustes distance between the centroids of each sex (vertical line) to a null distribution generated from 1000 permutations of the sex-labels for each of the six classification schemes. This result indicates that the difference between the centroids is as small/smaller than expected by random. For explination of classification scheme names see Table \ref{tab:hypotheses}.}
  \label{fig:sex_test_group}
\end{figure}

@@

@@

\begin{figure}[h]
  \centering
  \includegraphics[height = 0.5\textheight, width = \textwidth, keepaspectratio = true]{figure/other_model_sel}
  \caption{Comparisons of model fit to the training dataset for each of the supervised learning methods applied to the first two datasets; the results from the eight species dataset are presented in the left column, while those from the \textit{Trachemys} dataset are presented in the right column. Models were fit to datasets of varying complexity, with the number of parameters listed along the x-axis. Model fit is measured as the area under the receiver operating characteristic (AUC), which ranges from 0.5 to 1. Error bars correspond to one standard error estimated from 10 rounds of 5-fold cross-validation. The red dot corresponds to the model fit with the highest mean AUC while the blue dot corresponds to the model selected for further analysis. In some cases, there is no difference in complexity between the best and selected models.}
  \label{fig:other_sel}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[height = 0.3\textheight, width = \textwidth, keepaspectratio = true]{figure/other_oos_sel}
  \caption{The results of out-of-sample predictive performance of the selected models for both the eight species (left) and \textit{Trachemys} datasets. Predictive performance is measured as the area under the receiver operating characteristic (AUC), which ranges from 0.5 to 1. Points correspond to the individual out-of-sample predictive performance of the specific model, indicated along the x-axis. The horizontal bars correspond the average out-of-sample predictive performance of all the models.}
  \label{fig:other_oos}
\end{figure}

@@

\afterpage{\clearpage}
\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/emys_model_sel}
  \caption{AUC values for models of varying complexity fit to the \textit{Emys marmorata} training datasets for each classification scheme. The x-axis corresponds to the total number of predictors included in each model, while the y-axis corresponds to the AUC value which is a measure of goodness of fit for classification datasets. A model with a high AUC value corresponds to better classification performance than a model with a lower AUC value. Standard errors on AUC estimates are calculated from 10 rounds of 5-fold cross-validation. Indicated are the best performing and the selected models, in red and blue respectively. In some cases, there is no difference in complexity between the best and selected models.}
  \label{fig:emys_sel}
\end{figure}

\afterpage{\clearpage}
\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/emys_oos_sel}
  \caption{Comparison of out-of-sample AUC estimates from the predictions of selected models (Fig. \ref{fig:emys_sel}), grouped by classification scheme. The horizontal line in each panel corresponds to the average AUC value across all models of that classification scheme.}
  \label{fig:emys_oos}
\end{figure}

@@

@@

@@

\section*{Supporting information}

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
\paragraph*{S1 Fig.}
\label{S1_Fig}
{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).

>>>>

Three different geometric morphometric datasets describing turtle plastron variation were assembled for this analysis: 1) specimens from seven distinct emydine species and a deriochelyine outgroup (\textit{C. picta}); 2) \textit{T. scripta} specimens from the two main subspecies (\textit{T. scripta elegans} and \textit{T. scripta scripta}); and 3) \textit{E. marmorata} specimens from across the species' geographic range. The first two datasets are intended to serve as a test of whether machine learning techniques can differentiate species-level groupings of emydid turtles using plastron shape. We expect that the first case represents a low complexity dataset because of the high level of plastron shape disparity that exists among these species \cite{Claude2003a,Claude2006,Angielczyk2011}, whereas the second dataset should be relatively higher in complexity and more analogous to the \textit{E. marmorata} example. We predict that the \textit{E. marmorata} dataset should be of the highest complexity and our greatest challenge given the previous finding that only very subtle differences existed between geographically-distinct populations \cite{Holland1992}.

@@

\afterpage{\clearpage}
\begin{figure}[ht]
  \centering
  %\includegraphics[height = 0.8\textheight, width = \textwidth, keepaspectratio = true]{figure/Ken_Ang_SpLoc_final_comp}
  \caption{Geographic distribution of specimens sampled for comparing the hypothesized subdivisions of \textit{Emys marmorata}. Each hypothesized scheme has two or more possible classes. Sample size differs between schemes because of our ability to confidently assign museum specimens to the various schemes. The number of localities shown on each map is less than the number of specimens sampled because some localities produced multiple specimens. The different classification abbreviations are as follows: \textit{E. marmorata} = ``marm'', \textit{E. pallida} = ``pall'', Central Coast Ranges = ``CCR'', San Joaquin Valley = ``SJ,'' Baya Peninsula = ``Baja,'', and Sierra Foothills = ``Foothill.''}
  \label{fig:map}
\end{figure}

@@

Sex was known only for a subset of the total dataset and was not included as a predictor of classification. Instead, we estimated the degree to which specimens cluster morphologically by sex in order to determine how much of a potential biasing factor sexual dimorphism could be for our analysis of the \textit{E. marmorata} species complex (see below).

@@

\begin{figure}[ht]
  \centering
  %\includegraphics[height = 0.5\textheight, width = \textwidth, keepaspectratio = true]{figure/plastra}
  \caption{Depiction of general plastral shape of \textit{E. marmorata} and position of the 19 landmarks used in this study. Anterior is towards the top of the figure.}
  \label{fig:plastra}
\end{figure}

@@

Instead of relying on a single supervised learning method, we chose to use an ensemble approach where multiple model types are used in concert so that any congruence between them increases our support for that conclusion over another \cite{Hastie2009}. The supervised learning methods used here are named in Table \ref{tab:methods}. Each of these methods makes different assumptions, treats data differently, and can produce different classification results depending on the nature of the data \cite{Hastie2009}. For example, multinomial logistic regression is a type of generalized linear model, whereas random forest is itself an ensemble approach where multiple decision trees are fit to subsets of the full dataset and then averaged.

@@

The maximum set of possible predictors or features used for any model of our dataset is comprised of the first 25 principal components (PCs), scaled centroid size, the interaction between scaled centroid size and PC 1, and the interaction between scaled centroid size and PC 2. Additional interaction terms were not considered because of model complexity/sample size concerns. Size and the interaction between size and PCs 1 and 2 were included as predictors to account for known ontogenetic variation in plastron shape \cite{Angielczyk2013a} as well as potential size differences between classes, even if this is unlikely \cite{Seeliger1945,Holland1992}. These data constitute a ``maximum set'' because the best or selected models based on five-fold cross-validation need not, and likely will not, include all predictors possible (see below). Because our supervised learning models use PCs as predictors, this approach is in many ways analogous to PCA regression. PCA regression takes advantage of the reduction and orthogonality PCs to improve regression fit \cite{Hastie2009}. Because the PCs of shape are by definition orthogonal, they can easily serve as independent predictors or features of class membership without fear of collinearity.

We adopted a training and testing paradigm for selecting parsimonious models and estimating their overall error rates \cite{Hastie2009,Kuhn2013}. Within-sample model performance is inherently biased upwards, so model evaluation requires overcoming this bias. With very large sample sizes, as in this study, part of the sample can be used as the ``training set'' and the remainder acts as the ``testing set.'' In this approach, following all cleaning and vetting, the data are split into a training dataset and a testing dataset. The former is used for fitting the model whereas the later is used for measuring model performance, a process called model generalization. For each scheme, we limited the model training and testing to only those individuals with class labels for that scheme. In this analysis, we randomly divided 80\% of samples into the training set and the remaining 20\% into the testing set.

In classification studies, such as this one, a common metric of performance is the receiver operating characteristic (ROC) which is the relationship between the false and true positive rates \cite{Hastie2009}. The area under the ROC curve (AUC) is the derived estimate of model performance; AUC ranges from 0.5 to 1 which correspond to performance similar to random guesses and perfect classification rates, respectively \cite{Hastie2009}. Both ROC and AUC are preferable to simple classification accuracy when class membership is unbalanced, as it is in these analyses \cite{Hastie2009}. The standard ROC and AUC calculations are defined only for binary classifications, which is not the case for our eight species and \textit{Emys} complex datasets. To generalize this approach for situations with multiple response classes, we used an all-against-one strategy where the model AUC is the average of the AUC values from the multiple binary comparisons of one class compared to all others \cite{Hand2001}.

@@

@@

\begin{figure}[h]
  \centering
  %\includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/other_pc_graph}
  \caption{Two scatterplots of morphological differences from two of the three datasets analyzed in this study. (a) Scatterplot of the first two PCA axies from the landmarks from the eight different species dataset, and (b) the first two axes of variation from two subspecies of \textit{Trachemys} dataset. Point colors correspond to the categories within each dataset while point size is proportional to individual centroid size. In parentheses next to the axis labels are the percent of total variation accounted for by that axis. For both datasets there are clear distinctions between the different categories.}
  \label{fig:other_pca}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  %\includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/emys_pc_graph}
  \caption{Scatterplot of the first two axes of morphological variation in the \textit{Emys marmorata} dataset. Each panel corresponds to one of the six different classification schemes analyzed as part of this study (Tab. \ref{tab:hypotheses}). Point color corresponds to the categories within each scheme, and the class names correspond to geographic regions. Point size is proportional to centroid size of that specimen and the numbers in parentheses next to the axis labels are the percent of total variation accounted for along that axis.}
  \label{fig:emys_pca}
\end{figure}

@@

\begin{figure}[h]
  \centering
  %\includegraphics[height = 0.3\textheight, width = \textwidth, keepaspectratio = true]{figure/sex_test_hist}
  \caption{Comparison of observed Procrustes distance between the centroids of each sex (vertical line) to a null distribution generated from 1000 permutations of the sex-labels. This result indicates that the difference between the centroids is as small/smaller than expected by random.}
  \label{fig:sex_test}
\end{figure}

\begin{figure}[h]
  \centering
  %\includegraphics[height = 0.8\textheight, width = \textwidth, keepaspectratio = true]{figure/sex_test_hist_grouped}
  \caption{Comparison of observed Procrustes distance between the centroids of each sex (vertical line) to a null distribution generated from 1000 permutations of the sex-labels for each of the six classification schemes. This result indicates that the difference between the centroids is as small/smaller than expected by random. For explination of classification scheme names see Table \ref{tab:hypotheses}.}
  \label{fig:sex_test_group}
\end{figure}

@@

@@

\begin{figure}[h]
  \centering
  %\includegraphics[height = 0.5\textheight, width = \textwidth, keepaspectratio = true]{figure/other_model_sel}
  \caption{Comparisons of model fit to the training dataset for each of the supervised learning methods applied to the first two datasets; the results from the eight species dataset are presented in the left column, while those from the \textit{Trachemys} dataset are presented in the right column. Models were fit to datasets of varying complexity, with the number of parameters listed along the x-axis. Model fit is measured as the area under the receiver operating characteristic (AUC), which ranges from 0.5 to 1. Error bars correspond to one standard error estimated from 10 rounds of 5-fold cross-validation. The red dot corresponds to the model fit with the highest mean AUC while the blue dot corresponds to the model selected for further analysis. In some cases, there is no difference in complexity between the best and selected models.}
  \label{fig:other_sel}
\end{figure}

\begin{figure}[h]
  \centering
  %\includegraphics[height = 0.3\textheight, width = \textwidth, keepaspectratio = true]{figure/other_oos_sel}
  \caption{The results of out-of-sample predictive performance of the selected models for both the eight species (left) and \textit{Trachemys} datasets. Predictive performance is measured as the area under the receiver operating characteristic (AUC), which ranges from 0.5 to 1. Points correspond to the individual out-of-sample predictive performance of the specific model, indicated along the x-axis. The horizontal bars correspond the average out-of-sample predictive performance of all the models.}
  \label{fig:other_oos}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  %\includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/emys_model_sel}
  \caption{AUC values for models of varying complexity fit to the \textit{Emys marmorata} training datasets for each classification scheme. The x-axis corresponds to the total number of predictors included in each model, while the y-axis corresponds to the AUC value which is a measure of goodness of fit for classification datasets. A model with a high AUC value corresponds to better classification performance than a model with a lower AUC value. Standard errors on AUC estimates are calculated from 10 rounds of 5-fold cross-validation. Indicated are the best performing and the selected models, in red and blue respectively. In some cases, there is no difference in complexity between the best and selected models.}
  \label{fig:emys_sel}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  %\includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/emys_oos_sel}
  \caption{Comparison of out-of-sample AUC estimates from the predictions of selected models (Fig. \ref{fig:emys_sel}), grouped by classification scheme. The horizontal line in each panel corresponds to the average AUC value across all models of that classification scheme.}
  \label{fig:emys_oos}
\end{figure}

@@

%\section*{Supporting information}
%
%% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
%\paragraph*{S1 Fig.}
%\label{S1_Fig}
%{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).
